### 2015-01-11 mcmc.tex
autocorrelation function content

The autocorrelation function is defined by
\begin{equation}
C(\Delta) &\equiv& \frac{E[(\pars_t - \mu)\,(\pars_{t+\Delta} - \mu)]}{E[(\pars_t)^2]
\\
\mu &=& E[\pars_t]
\quad ;
\end{equation}
that is, it is the covariance of points separated by $\Delta$, taken
relative to the variance.
By construction
\begin{equation}
C(0) &=& 0
\\
\lim_{\Delta\rightarrow\infty} C(\Delta) &=& 0
\quad ,
\end{equation}
and by definition, the autocorrelation ``time'' $\tau$ is
\begin{equation}
\tau = \int C(\Delta)\,\dd\Delta
\quad .
\end{equation}
Heuristically, the autocorrelation time is the number of samples you
need to take before you get an independent sample.
More precisely, the autocorrelation time is the factor by which you
have to reduce the number of samples when you compute the variance on
a sampling-based expectation value estimate.\note{HOGG Explain more fully.}

Of course, the autocorrelation-time definition does not by itself provide
an \emph{estimator} for the time.

...Why is it is not a panacea? ...estimators ...getting stuck in one mode

...What plots should you make?
...Plots showing individual walkers going through the space.
...(now show plots that demonstrate a non-converged and a converged chain)...

>>>>

The autocorrelation function is defined by
\begin{eqnarray}
C(\Delta) &\equiv& \frac{E[(\pars_k - \mu)\,(\pars_{k+\Delta} - \mu)]}{E[(\pars - \mu)^2]}
\label{eq:autocorrelationfn}\\
\mu &=& E[\pars]
\quad ;
\end{eqnarray}
that is, it is the covariance of points separated by $\Delta$, taken
relative to the variance.
In the above equation, the assumption is that the samples $\pars_k$ are
given in the order that they were generated by the MCMC algorithm.
By construction
\begin{eqnarray}
C(0) &=& 1
\\
\lim_{\Delta\rightarrow\infty} C(\Delta) &=& 0
\quad ,
\end{eqnarray}
and by definition, the autocorrelation ``time'' $\tau$ is
\begin{eqnarray}
\tau = \sum_{\Delta = 0}^{\infty} C(\Delta)
\label{eq:autocorrelationtime}\quad .
\end{eqnarray}
Heuristically, the autocorrelation time is the number of samples you
need to take before you get an independent sample.
More precisely, the autocorrelation time is the factor by which you
have to reduce the number of samples when you compute the variance on
a sampling-based expectation value estimate.\note{HOGG Explain more fully.  Maybe also cite Fengji.}
It is slightly off-topic here, but related to this, if you have a long
chain of length $K$ and with autocorrelation time $\tau$ and you thin
the chain down to every $\tau$th sample (instead of using all $K$
samples), you will get just as good sampling-based estimates as if you
had used all $K$ samples.

A sampler with a better autocorrelation time is just better; you have
to do fewer $f(\pars)$ calls per independent sample, and you have to
run less time to get accurate sampling-based integral estimates.
A sampler that takes an independent sample every time would have an
autocorrelation time of unity, which is the best possible value; this
optimal sampling is only possible for problems where the sampling is
analytic (for example if the posterior is perfectly Gaussian with
known mean and variance).
In general, the best sampler will be different for different problems,
and we will (below) tune the samplers we have to do the best on the
problems we have; this tuning will also be problem-specific.
There are many heuristic bases on which different samplers might be
compared or tuned, but fundamentally it is lower autocorrelation time
that separates good samplers from bad ones and is the ultimate basis
on which we compare performance.\note{HOGG: Bash Dunkley and anyone
  else who has used something else to compare.}

Of course, the autocorrelation-time definition
(\ref{eq:autocorrelationtime}) does not by itself provide an
\emph{estimator} for the time.
Usually, we just take the naive estimator---the ratio of empirical
expectations implied by \equationname~(\ref{eq:autocorrelationfn}),
and then perform a naive sum to get the autocorrelation time.
However, almost all empirical estimators of the autocorrelation time
will end up being under-estimates (biased low), because the finite
length of the chain makes any subset seem more representative than it
truly is; indeed this is a sub-problem of the problem that convergence
can never be known empirically with good confidence.

Finally one last point about convergence:
Since all convergence tests are fundamentally heuristic, it is useful
to just make the heuristic visualization of the samples $\pars_k$ as a
function of $k$, in the order they were generated.
The chain is likely to be converged only if the random-walk process
crossed the domain of $\pars$ fully many times during the MCMC run.
Often some parameter directions are much worse than others; it is
worth looking at the chain in all parameter directions.
